Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.

### 2024-01-03 apples_and_oranges.qmd
changes to figures and tables

---
title: Apples and Oranges in Adolescent Depression Treatments.
subtitle: A conceptual and quantiative re-appraisal of control arms in clinical trials.
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note: 
abstract: |
  To follow
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
bibliography: emotion_concepts_paper.bib
format: 
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super

@@

Table 1 shows the results containing the means and standard errors for each level, derived as the average of 1000 metanalyses. 
```{r, warning=FALSE, message = FALSE, echo=FALSE}
# for some reason doesn't work, but will make work.
# tab_1_df <- df_mean_coefs_from_sim[, 1:length(df_mean_coefs_from_sim)-1]
# knitr:: kable(tab_1_df, digits = 2, 
#              col.names = c('Trial Arm', 'Beta', 'SE', 'Upper CI', 'Lower CI') )
```

Figure 1 graphs the results with the means of each level alongside their confidence intervals.

@@

ggplot(df_mean_coefs_from_sim, aes(x = mean_coefficient, y = 1:nrow(df_mean_coefs_from_sim),
                                   colour = condition, label = text_label)) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
  geom_text(vjust = +1.5, size = 4) +  # Adjust vjust and size as needed
  scale_size_continuous(guide = "none") +
  guides(colour = FALSE) + 
  scale_color_manual(values = c(redish_palette, blueish_palette)) +
  theme_minimal() +
  labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Type and Treatment Arm",
       subtitle = "metanalytically derived estimates of within group changes") +
  xlab("Standardized Mean Difference (SMD) with 95% CIs") +
  ylab("") +
  ylim(0, nrow(df_mean_coefs_from_sim) + 1) +
  xlim(-2.5, 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
  geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
  geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
  theme(axis.title.x = element_text(size = 14), 
        axis.text.x  = element_text(size = 14),
        axis.text.y = (element_blank()),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14))+
  geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
  geom_text(x = 0.2, y = 4.75, label = "Line of No Effect", color = "grey", vjust = 0.5, hjust = 0)

@@

As can be intuited from the confidence intervals displayed, there are statistically significant differences at p<0.05 between the two control arms. Table 2 illustrates this with the output of one of the simulated models.

@@

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}

knitr:: kable(coefficients_from_contrasts_medcontrol_psy_control, col.names = c("Beta", "SE", "Lower CI", "Upper CI"), digits = 2)
```

>>>>

---
title: Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.
subtitle: Re-visiting the relative efficacy Control Conditions..
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note: 
abstract: |
  To follow
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
bibliography: emotion_concepts_paper.bib
format: 
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super

@@

Figure 1 graphs the results with the means of each level alongside their confidence intervals. Table 1 shows the results containing the means and standard errors for each level, derived as the average of 1000 metanalyses. 
As can be intuited from the confidence intervals displayed, there are statistically significant differences at p<0.05 between the two control arms. Table 2 illustrates this with the output of one of the simulated models.

@@

plot_means <- ggplot(df_mean_coefs_from_sim, aes(x = mean_coefficient, y = 1:nrow(df_mean_coefs_from_sim),
                                   colour = condition, label = text_label)) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2, position = position_dodge(0.5)) +
  geom_text(vjust = +1.5, size = 4) +  # Adjust vjust and size as needed
  scale_size_continuous(guide = "none") +
  guides(colour = FALSE) + 
  scale_color_manual(values = c(redish_palette, blueish_palette)) +
  theme_minimal() +
  labs(x = "TE-random", y = NULL, title = "Adolescent Depression Trial Efficacy by Treatment Arm",
       subtitle = "metanalytically derived estimates of within group changes") +
  xlab("Standardized Mean Difference (SMD) with 95% CIs") +
  ylab("") +
  ylim(0, nrow(df_mean_coefs_from_sim) + 1) +
  xlim(-2.5, 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", size = 1.5, colour = "grey") +
  geom_segment(x = -1.25, xend = -1.75, y = 4.7, yend = 4.7, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.55, y = 4.9, label = "More Effective", color = "grey", vjust = 0.5, hjust = 1) +
  geom_segment(x = -1.35, xend = -0.85, y = 4.9, yend = 4.9, arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = -1.05, y = 5.05, label = "Less Effective", color = "grey", vjust = 0.5, hjust = 0)+
  theme(axis.title.x = element_text(size = 14), 
        axis.text.x  = element_text(size = 14),
        axis.text.y = (element_blank()),
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14))+
  geom_curve(aes(x = 0.18, y = 4.73, xend = 0.02, yend = 4.5), color = "grey", curvature = -0.2, arrow = arrow(length = unit(0.25, "cm"), type = "closed")) +
  geom_text(x = 0.2, y = 4.75, label = "Line of No Effect", color = "grey", vjust = 0.5, hjust = 0)

png("plotted_means.png")
plot_means
dev.off()

@@

![Fig 1. Means of the Four Trial Arms](plotted_means.png){width=50%}

@@

coefficients_from_contrasts_medcontrol_psy_control$zed_value <- coefficients_from_contrasts_medcontrol_psy_control$Beta/
coefficients_from_contrasts_medcontrol_psy_control$SE

coefficients_from_contrasts_medcontrol_psy_control$p_value <- format(1 - pnorm(coefficients_from_contrasts_medcontrol_psy_control$zed_value), scientific = TRUE)

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: Table 1
#| tbl-cap: Means for the four Trial Arms alongside 95% Confidence Intervals
tab_1_df <- df_mean_coefs_from_sim[, 1:length(df_mean_coefs_from_sim)-1]

knitr:: kable(tab_1_df, digits = 2)
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: Table 2
#| tbl-cap: Statistics of the Difference Between Medication and Psychotherapy Control Conditions
knitr:: kable(coefficients_from_contrasts_medcontrol_psy_control,col.names = c("Beta", "SE", "Lower CI", "Upper CI", "z-value", "p-value"), digits = 2)
